---
title: "RUVIII Test Script"
output: html_document
params: 
  k: 1
  clusters: c(10, 25)
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
# Need to be altered
load("/home/users/allstaff/pullin.j/RUV Evaluation/Data/Raw_analysis_without_beads.Rdata")
```

```{r, message = FALSE, warning = FALSE}
library(cytofkit)
library(RColorBrewer)
library(rsvd)
library(ruv)
library(pheatmap)
library(FlowSOM)
library(ConsensusClusterPlus)
library(flowCore)
library(CATALYST)
library(tidyverse)
```

```{r}
NUM_CLUSTERS <- 40
# take the k value from the input
k <- 1
color_clusters= c("#466791","#60bf37","#953ada","#4fbe6c","#ce49d3","#a7b43d","#5a51dc",
                  "#d49f36","#552095","#507f2d","#db37aa","#84b67c","#a06fda","#df462a","#5b83db",
                  "#c76c2d","#4f49a3","#82702d","#dd6bbb","#334c22","#d83979","#55baad","#dc4555",
                  "#62aad3","#8c3025","#417d61","#862977","#bba672","#403367","#da8a6d","#a79cd4",
                  "#71482c","#c689d0","#6b2940","#d593a7","#895c8b","#bd5975", "#00bfff", "#56ff0d",
                  "#ffff00")
```

### Data Preprocessing

#### Catalyst data-preprocessing
```{r}
# TODO
# Thoughts - this will require the fcs files of those files with beads that can be corerced into a flowest. 
# It will be neccesary to have the same donwsampling index for all of the data - but it would be difiuclt (impossible actually) to
# sample and then normalise - might have to refactor the sample out of tidyverse or use the the same trick we are already using
```

```{r}
# Notes:
# cell_clustering1 is the cluster
# sample_ids is the sample ids
# expr is the actual cells

all_data <- data.frame(sample = sample_ids, cluster = cell_clustering1, expr)

# Remember the median of markers
#raw_marker_medians <- all_data %>%
#  select(-sample, -cluster) %>% 
# summarise_all(median) %>% 
# as.matrix() %>% 
# as.vector()

# TODO Split the factor up somehow

# Subsample the data into only that which we are interested in
sub_data <- dplyr::filter(all_data, sample %in% c("1B1", "2B1", "3B1", "4B1", "5B1", "6B1"))

# TODO Automate this based on selection
NUM_SAMPLES <- 6

make_M <- function(clusters, norm_clus){
  M <- matrix(0, nrow = length(clusters), ncol = length(norm_clus))
  # rewrite the for loop
  for(i in 1:length(norm_clus)){
    M[clusters == norm_clus[i], i] = 1
  }
  as.matrix(M)
}

# Make the clusters
M <- make_M(sub_data$cluster, c(33,30,28)) 

# Append a column

sub_data <- cbind(sub_data, index = 1:nrow(sub_data))

sub_data <- sub_data %>% 
  group_by(sample) %>% 
  sample_n(10000) %>% 
  ungroup()

sub_data <- as.data.frame(sub_data)

M <- M[sub_data$index, ]

sub_data <- select(sub_data, -index)
```

#### PCA Plots based on a median column summary of the data

```{r}
#data <- all_data

#summ_data <- data %>% 
#  group_by(sample) %>% 
#  select(-cluster) %>% 
#  summarise_all(median)

#pca <- prcomp(as.matrix(summ_data[,-1]), scale. = FALSE, center = TRUE)

#plot_data <- data.frame(sample = summ_data[,1], pca$x[,1:3])

#ggplot(plot_data, aes(x = PC1, y = PC2, col = factor(sample))) + 
#  geom_point(alpha = 0.8) + 
#  labs(col = "Sample") + 
#  theme_bw()

#ggplot(plot_data, aes(x = PC1, y = PC3, col = factor(sample))) + 
#  geom_point(alpha = 0.8) + 
#  labs(col = "Sample") + 
#  theme_bw()

#ggplot(plot_data, aes(x = PC2, y = PC3, col = factor(sample))) + 
#  geom_point(alpha = 0.8) + 
#  labs(col = "Sample") + 
#  theme_bw()
```

```{r}
data <- sub_data
```

#### PCA Plots (Raw data)

```{r}
# Run PCA
N <- 2000
samp <- sample(1:nrow(data), N, replace = FALSE)

pca <- prcomp(data[samp, 3:ncol(data)], scale. = TRUE)$x
# Refore the data
pca_plot <- cbind(data[samp,1:2], pca)

# Plot the data
# PC1 vs PC2
ggplot(pca_plot, aes(x = PC1, y = PC2, col = sample)) + 
  geom_point(alpha = 0.3) + 
  labs(col = "Sample") + 
  theme_bw()

# PC2 vs PC3
ggplot(pca_plot, aes(x = PC2, y = PC3, col = sample)) + 
  geom_point(alpha = 0.3) + 
  labs(col = "Sample") + 
  theme_bw()

# PC1 vs PC3
ggplot(pca_plot, aes(x = PC1, y = PC3, col = sample)) + 
  geom_point(alpha = 0.3) + 
  labs(col = "Sample") + 
  theme_bw()
```

#### t-SNE Plot (Raw data)

```{r}
# Run TSNE
# Downsample 
samp_data <- sample_n(data, 5000, replace = FALSE)
tsne <- cytof_dimReduction(samp_data[3:ncol(samp_data)], method = "tsne")
# Rename and reform
colnames(tsne) <- c("tSNE1", "tSNE2")
tsne_plot <- cbind(samp_data[,1:2], tsne)
# Plot the data
ggplot(tsne_plot , aes(x = tSNE1, y = tSNE2, col = sample)) + 
  geom_point(alpha = 0.3) + 
  labs(col = "Sample") + 
  theme_bw()
```

#### Cluster Frequency Matrix (Raw data)

```{r}
# Find the frequency of each cluster/sample combination and store in matrix
clust_freq_mat <- matrix(0, nrow = NUM_CLUSTERS, ncol = NUM_SAMPLES)
num_sample <- rep(1:NUM_SAMPLES, each = 10000)
for(i in 1:nrow(data)) {
  clust_freq_mat[data$cluster[i], num_sample[i]] <- clust_freq_mat[data$cluster[i], num_sample[i]] + 1
}

clust_freq_mat <- 100 * apply(clust_freq_mat, 2, function(x) x/sum(x))

# Produce a marker matrix 
cell_amounts <- table(data$cluster)
cell_freqs <- round(cell_amounts/sum(cell_amounts) * 100, 2)
labels_row_percent <- paste0(1:NUM_CLUSTERS, " (", cell_freqs, "%)")
labels_row <- paste0(1:NUM_CLUSTERS)

# Reform the matrix into a tibble
plot_data <- tibble(
  centre = factor(rep(unique(data$sample), each = NUM_CLUSTERS), levels = unique(data$sample)),
  # Reverse the plotting layout
  cluster = factor(rep(labels_row_percent, NUM_SAMPLES), levels = rev(labels_row_percent)),
  freq = as.vector(clust_freq_mat)
)

# Plot the matrix
ggplot(plot_data, aes(factor(centre), factor(cluster))) + 
  geom_tile(aes(fill = freq), col = "black") + 
  theme_bw() + 
  labs(x = "Centre", y = "Cluster", fill = "Percentage of \ncells in cluster") + 
  scale_fill_gradientn(colours = (colorRampPalette(rev(brewer.pal(n = 11, name = "RdYlBu")))(50))) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_blank())
```

##### Histogram of between centre cluster amount variances

```{r}
logit <- function(p) p/(1-p)
clust_freq_mat <- logit(clust_freq_mat)
vars <- apply(clust_freq_mat, 1, function(x) var(x))
ggplot(as.tibble(vars), aes(value)) + 
  geom_histogram(colour = "steelblue", fill = "white") + 
  theme_bw()
```

##### Summary statistics of between centre cluster amount variances

```{r}
cat("Median of variances is: ", median(vars))
cat("IQR of variances is: ", IQR(vars))
```

##### Overall chi-squared statistic of between centre cluster amount variances

```{r}
chi_sq <- function(Y){
  Y_i. <- rowSums(Y)
  Y_.j <- colSums(Y)
  n <- sum(Y)
  stat <- 0 
  for(i in 1:nrow(Y)){
    for (j in 1:ncol(Y)){
       stat <- stat + (Y[i,j] - Y_i.[i]*Y_.j[j]/n)^2 / (Y_i.[i]*Y_.j[j]/n)
    }
  }
  stat
}

cat("Chi-squared statistic is:", chi_sq(clust_amount_mat))
```

#### Cluster Coloured t-SNE Plot (Raw data)

```{r}
# Cluster annotation tSNE
ggplot(tsne_plot , aes(x = tSNE1, y = tSNE2, col = factor(cluster))) + 
  geom_point(alpha = 0.6) + 
  labs(col = "Cluster") + 
  scale_color_manual(values = color_clusters) +
  theme_bw() + 
  # Don't really understand what this is doing
  guides(color = guide_legend(override.aes = list(size = 4), ncol = 2))
```

```{r}
sampled_data_markers <- select(samp_data, -sample, -cluster)

# And extract a single marker to colour by 
for (i in 1:length(sampled_data_markers)){
  col_marker <- sampled_data_markers[,i]
  marker_name <- colnames(sampled_data_markers)[i]
  plot_data <- data.frame(select(tsne_plot, -sample, -cluster), col_marker)
  plot <- ggplot(plot_data, aes(tSNE1, tSNE2, col = col_marker)) + 
    geom_point() + 
    # Colour code from Robinson workflow
    scale_colour_gradientn(colours = colorRampPalette(rev(brewer.pal(n = 11, name = "Spectral")))(50)) + 
    labs(x = "tSNE1", y = "tSNE2", color = marker_name) + 
    theme_bw()
  print(plot)
}
```

```{r}
# Add in some specualtive entripy of mixing
my_BatchEntropy <- function(dataset, batch0, L=100, M=100, k=100) {
#entropy of batch mixing
# L is the number bootstrapping times
# M is the number of randomly picked cells    
# k is the number of nearest neighbours of cell (from all batches) to check   
  
require(RANN)  
nbatches<-length(unique(batch0))

entropy<-matrix(0,L,1)
set.seed(42) 
for (boot in 1:L) {
  bootsamples<-sample(1:nrow(dataset),M)
  W21<-nn2(dataset,query=dataset[bootsamples,],k)
  
  for (i in 1:length(bootsamples)){
    
    for (j in 1:nbatches) {
    xi<-max(1,sum(batch0[W21$nn.idx[i,]]==j))
    entropy[boot]<-entropy[boot]+xi*log(xi)
    }
  }
}

return( (-1)*entropy/length(bootsamples) )
}

batch <- map_dbl(tsne_plot$sample, function(x) as.numeric(substr(x,1,1)))
test <- my_BatchEntropy(dataset = tsne_plot, batch0 = batch)
raw_entropy <- test
```


#### Median Marker Intensity Heatmap (Raw data)

```{r}
plot_data <- data %>% 
  group_by(cluster) %>%
  select(-sample) %>% 
  summarise_all(median) %>% 
  gather(key = cluster) %>% 
  bind_cols(cluster = factor(rep(1:NUM_CLUSTERS, 14), levels = as.character(1:NUM_CLUSTERS))) %>% # This is a bit messy
  set_names(c("marker", "intensity", "cluster"))

plot <- ggplot(plot_data, aes(factor(marker), factor(cluster))) + 
  geom_tile(aes(fill = intensity), col = "black") + 
  theme_bw() + 
  labs(x = "Cluster", y = "Marker", fill = "Median Intensity") + 
  scale_fill_gradientn(colours = (colorRampPalette(rev(brewer.pal(n = 11, name = "RdYlBu")))(50))) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_blank())
plot
```

#### F Statistics (Raw Data)

##### F Statsitics based on marker measurements

###### Heatmap of F statistics
```{r}
F_stats <- matrix(0, nrow = NUM_CLUSTERS, ncol = length(3:ncol(data)))
for (j in 1:NUM_CLUSTERS){
  cluster_data <- dplyr::filter(data, cluster == j)
  for(i in 3:ncol(data)) {
    marker <- cluster_data[ ,i]
    anova_data <- tibble(sample = cluster_data$sample, marker)
    model <- aov(marker ~ sample, anova_data)
    F_stats[j,i-2] <- summary(model)[[1]]["F value"][[1]][[1]]
  }
}

F_stats <- log(F_stats)

pheatmap(F_stats, 
         cluster_rows = FALSE, 
         cluster_cols = FALSE, 
         labels_col = colnames(data)[3:ncol(data)], 
         labels_row = 1:NUM_CLUSTERS)
```

###### Histogram summary of F Statistics

```{r}
ggplot(as.tibble(as.vector(F_stats)), aes(value)) + 
  geom_histogram(colour = "steelblue", fill = "white", binwidth = 0.5) + 
  theme_bw()
```

##### F Statistics based on PCA varaibles

###### Heatmap of F statistics
```{r}
NUM_PCS <- 5 

pca_data <- NULL
pca_data <- prcomp(data[,3:ncol(data)], scale. = FALSE, center = FALSE)$x[,1:NUM_PCS]
pca_data <- as.data.frame(cbind(data$cluster, data$sample, pca_data))
colnames(pca_data)[1:2] <- c("cluster", "sample")

F_stats <- matrix(0, nrow = NUM_CLUSTERS, ncol = NUM_PCS)
for (j in 1:NUM_CLUSTERS){
  cluster_data <- dplyr::filter(pca_data, cluster == j)
  for(i in 3:ncol(pca_data)) {
    prin_comp <- cluster_data[ ,i]
    anova_data <- tibble(sample = cluster_data$sample, prin_comp)
    model <- aov(prin_comp ~ sample, anova_data)
    F_stats[j,i-2] <- summary(model)[[1]]["F value"][[1]][[1]]
  }
}

F_stats <- log(F_stats + 1)

pheatmap(F_stats, 
         cluster_rows = FALSE, 
         cluster_cols = FALSE, 
         labels_col = colnames(pca_data)[3:ncol(pca_data)], 
         labels_row = 1:NUM_CLUSTERS)
```

###### Histogram summary of F Statistics

```{r}
ggplot(as.tibble(as.vector(F_stats)), aes(value)) + 
  geom_histogram(colour = "steelblue", fill = "white", binwidth = 0.5) + 
  theme_bw()
```

##### F Statistics based on tSNE varaibles

###### Heatmap of F statistics
```{r}
F_stats <- matrix(0, nrow = NUM_CLUSTERS, ncol = 2)
for (j in 1:NUM_CLUSTERS){
  cluster_data <- dplyr::filter(tsne_plot, cluster == j)
  for(i in 3:ncol(tsne_plot)) {
    if(nrow(cluster_data) > 10){
      tsne_dim <- cluster_data[ ,i]
      anova_data <- tibble(sample = cluster_data$sample, tsne_dim)
      model <- aov(tsne_dim ~ sample, anova_data)
      F_stats[j,i-2] <- summary(model)[[1]]["F value"][[1]][[1]]
    } else {
      F_stats[j,i-2] <- NA
    }
  }
}

F_stats <- log(F_stats + 1)

pheatmap(F_stats, 
         cluster_rows = FALSE, 
         cluster_cols = FALSE, 
         labels_col = colnames(tsne_plot)[3:ncol(tsne_plot)], 
         labels_row = 1:NUM_CLUSTERS)

F_stats <- as.vector(F_stats)[!is.na(F_stats)]
raw_tsne_F_stat_range <- as.vector(range(F_stats))
```

###### Histogram summary of F Statistics

```{r}
ggplot(as.tibble(as.vector(F_stats)), aes(value)) + 
  geom_histogram(colour = "black", fill = "skyblue", binwidth = 0.25) + 
  theme_bw() + 
  geom_vline(xintercept = median(as.vector(F_stats)), col = "red") + 
  labs(y = "Count", x = "F-stats")
```

##### Between cluster F Statistics based on markers 
###### Heatmap of F statistics
```{r}
#sample_labels <- c("1A1", "2A2", "3A2", "4A2", "5A2", "6A2")

#F_stats <- matrix(0, nrow = NUM_SAMPLES, ncol = length(3:ncol(data)))
#for (j in 1:NUM_SAMPLES){
#  sample_data <- filter(data, as.character(sample) == sample_labels[j])
#  for(i in 3:ncol(data)) {
#    marker <- sample_data[ ,i]
#    anova_data <- tibble(cluster = sample_data$cluster, marker)
#    head(anova_data)
#    model <- aov(marker ~ cluster, anova_data)
#    F_stats[j,i-2] <- summary(model)[[1]]["F value"][[1]][[1]]
#  }
#}

#pheatmap(F_stats, 
#         cluster_rows = FALSE,
#        cluster_cols = FALSE, 
#         labels_col = colnames(data)[3:ncol(data)], 
#         labels_row = sample_labels)
```

###### Histogram summary of F Statistics

```{r}
#ggplot(as.tibble(as.vector(F_stats)), aes(value)) + 
#  geom_histogram(colour = "steelblue", fill = "white") + 
#  theme_bw()
```

## Run RUVIII and recluster the data

```{r}
Y <- as.matrix(select(sub_data, -sample, -cluster))

fastRUVIII = function(Y, M, ctl, k=NULL, eta=NULL, average=FALSE, fullalpha=NULL){
  # Assumes good input
  Y = RUV1(Y,eta,ctl)
  m = nrow(Y)
  Y0 = residop(Y, M)
  fullalpha = diag(rsvd(Y0)$d) %*% t(rsvd(Y0)$v)
  alpha = fullalpha[1:k,,drop=FALSE]
  ac = alpha[,ctl,drop=FALSE]
  W = Y[,ctl] %*% t(ac) %*% solve(ac %*% t(ac))
  newY = Y - W %*% alpha
  return(list(newY = newY, fullalpha=fullalpha))
}

# Take all the columns as negative controls
# Take eta to be 1
norm_Y <- fastRUVIII(Y, M, ctl = c(1:ncol(Y)), eta = 1, k = k)$newY

# TODO
# Add back the median of the markers before normalisation 

out <- data.frame(sample = sub_data$sample, cluster = sub_data$cluster, norm_Y) 
```

```{r, message=FALSE}
# Need to recluster the data
set.seed(1234)

new_ff <- fcs[[1]]

new_ff@exprs <- norm_Y

fsom <- ReadInput(new_ff, transform = FALSE, scale = FALSE)
som <- BuildSOM(fsom, colsToUse = c(lineage_markers))
cell_clustering_som <- som$map$mapping[,1]

## Metaclustering into NUM_CLUSTERS clusters with ConsensusClusterPlus
codes <- som$map$codes
nmc <- NUM_CLUSTERS
mc <- ConsensusClusterPlus(t(codes), maxK = nmc, reps = 100, 
                           pItem = 0.9, pFeature = 1, title = plot_outdir, plot = "png", 
                           clusterAlg = "hc", innerLinkage = "average", finalLinkage = "average", 
                           distance = "euclidean", seed = 1234)

## Get cluster ids for each cell
new_clusters <- code_clustering1[som$map$mapping[,1]]
out$cluster <- new_clusters
```

```{r}
data <- out
```

#### PCA Plots (Normalised data)

```{r}
# Run PCA
N <- 2000
samp <- sample(1:nrow(data), N, replace = FALSE)

pca <- prcomp(data[samp, 3:ncol(data)], scale. = TRUE)$x
# Refore the data
pca_plot <- cbind(data[samp,1:2], pca)

# Plot the data
# PC1 vs PC2
ggplot(pca_plot, aes(x = PC1, y = PC2, col = sample)) + 
  geom_point(alpha = 0.3) + 
  labs(col = "Sample") + 
  theme_bw()

# PC2 vs PC3
ggplot(pca_plot, aes(x = PC2, y = PC3, col = sample)) + 
  geom_point(alpha = 0.3) + 
  labs(col = "Sample") + 
  theme_bw()

# PC1 vs PC3
ggplot(pca_plot, aes(x = PC1, y = PC3, col = sample)) + 
  geom_point(alpha = 0.3) + 
  labs(col = "Sample") + 
  theme_bw()
```

#### t-SNE Plot Coloured by Sample (Normalised data)

```{r}
# Run TSNE
# Downsample 
samp_data <- sample_n(data, 5000, replace = FALSE)
tsne <- cytof_dimReduction(samp_data[3:ncol(samp_data)], method = "tsne")
# Rename and reform
colnames(tsne) <- c("tSNE1", "tSNE2")
tsne_plot <- cbind(samp_data[,1:2], tsne)
# Plot the data
ggplot(tsne_plot , aes(x = tSNE1, y = tSNE2, col = sample)) + 
  geom_point(alpha = 0.3) + 
  labs(col = "Sample") + 
  theme_bw()
```

#### Cluster Frequency Matrix (Normalised data)

```{r}
# Find the frequency of each cluster/sample combination and store in matrix
clust_freq_mat <- matrix(0, nrow = NUM_CLUSTERS, ncol = NUM_SAMPLES)
num_sample <- rep(1:NUM_SAMPLES, each = 10000)
for(i in 1:nrow(data)) {
  clust_freq_mat[data$cluster[i], num_sample[i]] <- clust_freq_mat[data$cluster[i], num_sample[i]] + 1
}

clust_freq_mat <- 100 * apply(clust_freq_mat, 2, function(x) x/sum(x))

# Produce a marker matrix 
cell_amounts <- table(data$cluster)
cell_freqs <- round(cell_amounts/sum(cell_amounts) * 100, 2)
labels_row_percent <- paste0(1:NUM_CLUSTERS, " (", cell_freqs, "%)")
labels_row <- paste0(1:NUM_CLUSTERS)

# Reform the matrix into a tibble
plot_data <- tibble(
  centre = factor(rep(unique(data$sample), each = NUM_CLUSTERS), levels = unique(data$sample)),
  # Reverse the plotting layout
  cluster = factor(rep(labels_row_percent, NUM_SAMPLES), levels = rev(labels_row_percent)),
  freq = as.vector(clust_freq_mat)
)

# Plot the matrix
ggplot(plot_data, aes(factor(centre), factor(cluster))) + 
  geom_tile(aes(fill = freq), col = "black") + 
  theme_bw() + 
  labs(x = "Centre", y = "Cluster", fill = "Percentage of \ncells in cluster") + 
  scale_fill_gradientn(colours = (colorRampPalette(rev(brewer.pal(n = 11, name = "RdYlBu")))(50))) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_blank())
```

##### Histogram of between centre cluster amount variances

```{r}
clust_amount_mat <- asinh(clust_freq_mat/5)
vars <- apply(clust_amount_mat, 1, function(x) var(x))
ggplot(as.tibble(vars), aes(value)) + 
  geom_histogram(colour = "steelblue", fill = "white", binwidth = 0.005) + 
  theme_bw()

```

##### Summary statistics of between centre cluster amount variances

```{r}
cat("Median of variances is: ", median(vars))
cat("IQR of variances is: ", IQR(vars))
```

##### Overall chi-squared statistic of between centre cluster amount variances

```{r}
chi_sq <- function(Y){
  Y_i. <- rowSums(Y)
  Y_.j <- colSums(Y)
  n <- sum(Y)
  stat <- 0 
  for(i in 1:nrow(Y)){
    for (j in 1:ncol(Y)){
       stat <- stat + (Y[i,j] - Y_i.[i]*Y_.j[j]/n)^2 / (Y_i.[i]*Y_.j[j]/n)
    }
  }
  stat
}

cat("Chi-squared statistic is:", chi_sq(clust_amount_mat))
```

#### Cluster Coloured t-SNE Plot (Normalised data)

```{r}
# Cluster annotation tSNE
ggplot(tsne_plot , aes(x = tSNE1, y = tSNE2, col = factor(cluster))) + 
  geom_point(alpha = 0.6) + 
  labs(col = "Cluster") + 
  scale_color_manual(values = color_clusters) +
  theme_bw() + 
  # Don't really understand what this is doing
  guides(color = guide_legend(override.aes = list(size = 4), ncol = 2))
```

```{r}
sampled_data_markers <- select(samp_data, -sample, -cluster)

# And extract a single marker to colour by 
for (i in 1:length(sampled_data_markers)){
  col_marker <- sampled_data_markers[,i]
  marker_name <- colnames(sampled_data_markers)[i]
  plot_data <- data.frame(select(tsne_plot, -sample, -cluster), col_marker)
  plot <- ggplot(plot_data, aes(tSNE1, tSNE2, col = col_marker)) + 
    geom_point() + 
    # Colour code from Robinson workflow
    scale_colour_gradientn(colours = colorRampPalette(rev(brewer.pal(n = 11, name = "Spectral")))(50)) + 
    labs(x = "tSNE1", y = "tSNE2", color = marker_name) + 
    theme_bw()
  print(plot)
}
```


```{r}
batch <- map_dbl(tsne_plot$sample, function(x) as.numeric(substr(x,1,1)))
test <- my_BatchEntropy(dataset = tsne_plot, batch0 = batch)
norm_entropy <- test
```

#### Median Marker Intensity Heatmap (Normalised data)

```{r}
plot_data <- data %>% 
  group_by(cluster) %>%
  select(-sample) %>% 
  summarise_all(median) %>% 
  gather(key = cluster) %>% 
  bind_cols(cluster = factor(rep(1:NUM_CLUSTERS, 14), levels = as.character(1:NUM_CLUSTERS))) %>% # This is a bit messy
  set_names(c("marker", "intensity", "cluster"))


plot <- ggplot(plot_data, aes(factor(marker), factor(cluster))) + 
  geom_tile(aes(fill = intensity), col = "black") + 
  theme_bw() + 
  labs(x = "Cluster", y = "Marker", fill = "Median Intensity") + 
  scale_fill_gradientn(colours = (colorRampPalette(rev(brewer.pal(n = 11, name = "RdYlBu")))(50))) + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_blank())
plot
```

#### F Statistics (Normalised Data)
##### F Statsitics based on marker measurements

###### Heatmap of F statistics
```{r}
F_stats <- matrix(0, nrow = NUM_CLUSTERS, ncol = length(3:ncol(data)))
for (j in 1:NUM_CLUSTERS){
  cluster_data <- dplyr::filter(data, cluster == j)
  for(i in 3:ncol(data)) {
    marker <- cluster_data[ ,i]
    anova_data <- tibble(sample = cluster_data$sample, marker)
    model <- aov(marker ~ sample, anova_data)
    F_stats[j,i-2] <- summary(model)[[1]]["F value"][[1]][[1]]
  }
}

F_stats <- log(F_stats)

pheatmap(F_stats, 
         cluster_rows = FALSE, 
         cluster_cols = FALSE, 
         labels_col = colnames(data)[3:ncol(data)], 
         labels_row = 1:NUM_CLUSTERS)
```

###### Histogram summary of F Statistics

```{r}
ggplot(as.tibble(as.vector(F_stats)), aes(value)) + 
  geom_histogram(colour = "steelblue", fill = "white", binwidth = 0.5) + 
  theme_bw()
```

##### F Statistics based on PCA variables

###### Heatmap of F statistics
```{r}
NUM_PCS <- 5 

pca_data <- NULL
pca_data <- prcomp(data[,3:ncol(data)], scale. = FALSE, center = FALSE)$x[,1:NUM_PCS]
pca_data <- as.data.frame(cbind(data$cluster, data$sample, pca_data))
colnames(pca_data)[1:2] <- c("cluster", "sample")

F_stats <- matrix(0, nrow = NUM_CLUSTERS, ncol = NUM_PCS)
for (j in 1:NUM_CLUSTERS){
  cluster_data <- dplyr::filter(pca_data, cluster == j)
  for(i in 3:ncol(pca_data)) {
    prin_comp <- cluster_data[ ,i]
    anova_data <- tibble(sample = cluster_data$sample, prin_comp)
    model <- aov(prin_comp ~ sample, anova_data)
    F_stats[j,i-2] <- summary(model)[[1]]["F value"][[1]][[1]]
  }
}

F_stats <- log(F_stats + 1)

pheatmap(F_stats, 
         cluster_rows = FALSE, 
         cluster_cols = FALSE, 
         labels_col = colnames(pca_data)[3:ncol(pca_data)], 
         labels_row = 1:NUM_CLUSTERS)
```

###### Histogram summary of F Statistics

```{r}
ggplot(as.tibble(as.vector(F_stats)), aes(value)) + 
  geom_histogram(colour = "steelblue", fill = "white", binwidth = 0.5) + 
  theme_bw()
```

##### F Statistics based on tSNE1 varaibles

###### Heatmap of F statistics
```{r}
F_stats <- matrix(0, nrow = NUM_CLUSTERS, ncol = 2)
for (j in 1:NUM_CLUSTERS){
  cluster_data <- dplyr::filter(tsne_plot, cluster == j)
  for(i in 3:ncol(tsne_plot)) {
    if(nrow(cluster_data) > 10){
      tsne_dim <- cluster_data[ ,i]
      anova_data <- tibble(sample = cluster_data$sample, tsne_dim)
      model <- aov(tsne_dim ~ sample, anova_data)
      F_stats[j,i-2] <- summary(model)[[1]]["F value"][[1]][[1]]
    } else {
      F_stats[j,i-2] <- NA
    }
  }
}

F_stats <- log(F_stats + 1)

pheatmap(F_stats, 
         cluster_rows = FALSE, 
         cluster_cols = FALSE, 
         labels_col = colnames(tsne_plot)[3:ncol(tsne_plot)], 
         labels_row = 1:NUM_CLUSTERS)

F_stats <- as.vector(F_stats)[!is.na(F_stats)]
raw_tsne_F_stat_range <- as.vector(range(F_stats))
```

###### Histogram summary of F Statistics

```{r}
ggplot(as.tibble(as.vector(F_stats)), aes(value)) + 
  geom_histogram(colour = "black", fill = "skyblue", binwidth = 0.25) + 
  theme_bw() + 
  geom_vline(xintercept = median(as.vector(F_stats)), col = "red") + 
  labs(y = "Count", x = "F-stats")
```

##### Between cluster F Statistics based on markers 

###### Heatmap of F statistics
```{r}
#sample_labels <- c("1A1", "2A2", "3A2", "4A2", "5A2", "6A2")

#F_stats <- matrix(0, nrow = NUM_SAMPLES, ncol = length(3:ncol(data)))
#for (j in 1:NUM_SAMPLES){
#  sample_data <- filter(data, as.character(sample) == sample_labels[j])
#  for(i in 3:ncol(data)) {
#    marker <- sample_data[ ,i]
#    anova_data <- tibble(cluster = sample_data$cluster, marker)
#    head(anova_data)
#    model <- aov(marker ~ cluster, anova_data)
#    F_stats[j,i-2] <- summary(model)[[1]]["F value"][[1]][[1]]
#  }
#}

#pheatmap(F_stats, 
#         cluster_rows = FALSE,
#         cluster_cols = FALSE, 
#         labels_col = colnames(data)[3:ncol(data)], 
#         labels_row = sample_labels)
```

###### Histogram summary of F Statistics

```{r}
#ggplot(as.tibble(as.vector(F_stats)), aes(value)) + 
#  geom_histogram(colour = "steelblue", fill = "white") + 
#  theme_bw()
```

###### Entropy boxplot

```{r}
entropy_data <- tibble(type = c(rep("Raw Data", 100), rep("RUVIII Nomrmalisation", 100)),
                       entropy = c(raw_entropy, norm_entropy))

ggplot(entropy_data, aes(x = factor(type), y = entropy)) + 
  geom_boxplot() + 
  labs(x = "", y = "Entropy of mixing") + 
  theme_bw()
```

#### Cluster Matching Heatmap

```{r}
freq_mat <- matrix(0, nrow = NUM_CLUSTERS, ncol = NUM_CLUSTERS)
for (i in 1:nrow(sub_data)){
  freq_mat[sub_data$cluster[i], out$cluster[i]] <- freq_mat[sub_data$cluster[i], out$cluster[i]] + 1
}

# Log taken to ensure that no single observation domimates the colour scheme
freq_mat <- log(freq_mat + 1)

pheatmap(freq_mat, 
         cluster_rows = FALSE, 
         cluster_cols = FALSE)
```

```{r}
# TODO 
# add canonical correlation analysis
```

